---
title: "Capstone 1"
author: "Curtis"
date: "2023-08-13"
output: html_document
---

```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
```

Importing the 12 months CSV 
```{r Importing}
raw202207 <- read_csv("Raw Data/202207-divvy-tripdata.csv")
raw202208 <- read_csv("Raw Data/202208-divvy-tripdata.csv")
raw202209 <- read_csv("Raw Data/202209-divvy-publictripdata.csv")
raw202210 <- read_csv("Raw Data/202210-divvy-tripdata.csv")
raw202211 <- read_csv("Raw Data/202211-divvy-tripdata.csv")
raw202212 <- read_csv("Raw Data/202212-divvy-tripdata.csv")
raw202301 <- read_csv("Raw Data/202301-divvy-tripdata.csv")
raw202302 <- read_csv("Raw Data/202302-divvy-tripdata.csv")
raw202303 <- read_csv("Raw Data/202303-divvy-tripdata.csv")
raw202304 <- read_csv("Raw Data/202304-divvy-tripdata.csv")
raw202305 <- read_csv("Raw Data/202305-divvy-tripdata.csv")
raw202306 <- read_csv("Raw Data/202306-divvy-tripdata.csv")

```
Checking the column names if there's any differences
```{r Checking Column}
colnames(raw202207)
colnames(raw202208)
colnames(raw202209)
colnames(raw202210)
colnames(raw202211)
colnames(raw202212)
colnames(raw202301)
colnames(raw202302)
colnames(raw202303)
colnames(raw202304)
colnames(raw202305)
colnames(raw202306)
```

```{r Checking Str}
str(raw202207)
str(raw202208)
str(raw202209)
str(raw202210)
str(raw202211)
str(raw202212)
str(raw202301)
str(raw202302)
str(raw202303)
str(raw202304)
str(raw202305)
str(raw202306)
```

Stacking Individual Month's into one big Data Frame
```{r Stacking Into Big Data Frame}
all_trips <- bind_rows(raw202207, raw202208, raw202209, raw202210, raw202211,  raw202212, raw202301, raw202302, raw202303, raw202304, raw202305, raw202306)
```

```{r}
rm(raw202207)
rm(raw202208)
rm(raw202209)
rm(raw202210)
rm(raw202211)
rm(raw202212)
rm(raw202301)
rm(raw202302)
rm(raw202303)
rm(raw202304)
rm(raw202305)
rm(raw202306)
```


removing start_lat, start_lng, end_lat, end_lng, birthyear
```{r Removing Unneeded columns}
all_trips <- all_trips %>% 
  select(-c(start_lat, start_lng, end_lat, end_lng))
```

Checking for the new dataframe
```{r}
  colnames(all_trips)
  nrow(all_trips)
  dim(all_trips)
  head(all_trips)
  tail(all_trips)
  str(all_trips)
  summary(all_trips)
```
Adding columns that list the date, month, day, and year of each ride.
```{r}
all_trips$date <- as.Date(all_trips$started_at)
all_trips$month <- format(as.Date(all_trips$date), "%m")
all_trips$day <- format(as.Date(all_trips$date), "%d")
all_trips$year <- format(as.Date(all_trips$date), "%Y")
all_trips$day_of_week <- format(as.Date(all_trips$date), "%A")
```

Checking the added columns
```{r}
head(all_trips)
```

Adding a "ride_length" calculation by subtracting "ended_at" with "started_at"

```{r}
all_trips$ride_length <- difftime(all_trips$ended_at, all_trips$started_at)
```

```{r}
str(all_trips)
```

converting "ride_length" from Factor to numeric so we can run calculations on the data
```{r}
is.factor(all_trips$ride_length)
all_trips$ride_length <- as.numeric(as.character(all_trips$ride_length))
is.numeric(all_trips$ride_length)
```
Creating new data frame where we remove the bike that requires maintenance
```{r}
all_trips_v2 <- all_trips[!(all_trips$start_station_name == 'HQ QR' | all_trips$ride_length<0),]
```
descriptive analysis on ride_length

```{r}
summary(all_trips_v2$ride_length)
```

```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = mean)
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = median)
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = max)
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = min)
```
```{r}
all_trips_v2$day_of_week <- ordered(all_trips_v2$day_of_week, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
```


average ride time by each day for members vs casual users
```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual + all_trips_v2$day_of_week, FUN = mean)
```
analyzing ridership data by type and weekday
```{r}
all_trips_v2 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  summarize(number_of_rides = n(), average_duration = mean(ride_length)) %>% 
  arrange(member_casual, weekday)
```

Visualizing the number of rides by rider type
```{r}
all_trips_v2 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  summarize(number_of_rides = n(), average_duration = mean(ride_length)) %>% 
  arrange(member_casual, weekday) %>% 
  ggplot(aes(x = weekday, y = number_of_rides, fill=member_casual))+
  geom_col(position="dodge")
```
```{r}
all_trips_v2 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  summarize(number_of_rides = n(), average_duration = mean(ride_length)) %>% 
  arrange(member_casual, weekday) %>% 
  ggplot(aes(x = weekday, y = average_duration, fill=member_casual))+
  geom_col(position="dodge")
```

```{r}
counts <- aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual + all_trips_v2$day_of_week, FUN = mean)

write.csv(counts, file = 'D:/Study/R Work Directory/Capstone/all_trips_v2done.csv')
```

